Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends

### 2021-12-26 Paper_online_Learning.tex
update introduction

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
            pdfkeywords={Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

@@

\begin{document}
\maketitle
\begin{abstract}
Distance Education has long history. However Coid-19 has made a new era of distance education. With the increasing demand, many different distance learning solutions have been introduced for different education purposes. In this study we invetigated the impact of Covid 19 on ditance education and the possibility of using Google analytics search index series as a proxy to quantify the popularity and the public interest for different distance education solutions. Both visual and analytical approaches were used to analyse web search queries at the global level during Covid-19 pandemic.This study intorduced a Google trend footprint of various ditance leraning solutions for diffrent education purposes, that can be used as a proxy for measuring global attension towards different learning solutions. The findings provide a fast first step to identify the most popular tools available for different educational purposes. R code and data are available in the online supplementary materials.
\end{abstract}
\begin{keywords}
Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries
\end{keywords}

@@

The ongoing pandemic of COVID-19 is still one of the most important priorities of governments and media of many countries all around the world. Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}. As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}. In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.

@@

The ongoing COVID-19 pandemic has brought the world to a standstill. Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 \autocite{unesco2020covid}. With the start of the pandemic a near total closure of schools were observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools where classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.

The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders: teachers, students and parents. Teachers were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic. Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services, long-term educational disengagement, digital exclusion, poor technology management, increased psychosocial challenges, decreases in mathematical achievement, a lack of engagement, increased early marriages , psychosocial challenges \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}. This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home. Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures. Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown. This situation left internet as the only medium to support all these educational processes.

In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}. However, getting familiar with all these available distance solutions is not feasible or practical due to limited time availability and need of urgent responses. According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.

Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting distance education solutions. Further, according to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as popularity can indicate the prominence of the product in the market, its usability and its impact. Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search. Therefore, the internet represents a great opportunity to learn about public attention towards a product or service. It is also a way to narrow down the search space and thereby identify most suitable product and services that meets customer needs.

Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}. As there is still the main struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic? (3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021. The resulted google trend footprint provides a fast first step to identify the most popular distance education tools available for different education purposes. It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching. Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.

Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal. Having the right tool is imperative to successful completion of a task at hand. The findings of this study provides an initial guidance to select a right tool. Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.

@@

To identify the impact of covid 19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO \autocite{unesco2020covid}. Data related to Covid-19 cases was retrieved from the R software \autocite{rsoftware} package \texttt{coronavirus} \autocite{coronavirusr}. The package pulls data from the Johns
Hopkins University Center for Systems Science and Engineering (JHU CCSE) data repository \autocite{dong2020interactive}.

In this work we also used Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions. We used weekly Google Analytics search index series of various distance learning solutions during covid-19 pandemic. We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of Covid-19 pandemic. Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time. We also studied the cross correlation between weekly web searches and the weekly cases of covid-19.

@@

Google Trends allows to search for up to five queries each time \autocite{vaughan2014web}. We did our analysis under 11 sub segmentations namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support. The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to Covid 19 pandemic.

@@

Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the Covid-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education. We also evaluated the relationship between weekly web searches and the number of weekly global cases of covid-19. Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of covid-19 related search volume data and distance education related search volume data.

@@

Search volume index graphs for Covid 19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic. Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic. According to \textcite{moore2011learning}, ``distance education'' is the most renowned descriptor used when referencing distance learning. However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``never version'' and ``improved version'' of distance learning. The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout Covid-19 pandemic. Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period. Covid-19 related web searches declined in the months after the WHO announcement of the Covid-19 pandemic, despite the increase in Covid-19 cases. A similar decline could be detected in distance education related terms. further, a second was observed only in education related terms after the WHO announcement of the Covid-19 pandemic. This could be due to the near total closures of schools happened with the declaration of Covid-19 as a pandemic.

@@

The pairwise comparison performed using Dynamic Time Warping algorithm. Under this approach the two time series are stretched or compressed locally in order to make one resemble the other as much as possible. The distance between the two series is then computed by summing the distances of individual aligned elements \autocite{giorgino2009computing}. Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from covid 19 on distance education. Both pairwise comparison show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)). The normalized distance between the two series in the two comparisons are 0.1326197 and 0.1828299, respectively. This also provides further confirmation for the similarity of the google trend search profiles.

The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to covid 19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after covid 19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).

The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide. By the time it was know as ``corona'', it was not a global concern and was mostly limited to China. During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to Covid 19 pandemic all around the world. This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis.

Later with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020. With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}). With this sudden unexpected movement, online learning also became an equally important global concern, similar to covid 19 pandemic. This similar behaviour in public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line from approximately 120 to 150 on the Query Index axis. This also confirms the impact of covid 19 on distance education and online learning.

Then, a cross correlation analysis was performed to estimate the lag between covid 19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}). A close relationship was evident between covid related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''. The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the covid related search terms and distance learning related search terms. The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)). However, no significant synchronization was observed between the Covid 19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).

@@

\caption{Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic}\label{fig:ccfAnalysis}
\end{figure}

@@

Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during covid 19 pandemic (Figure \ref{fig:plot1} (a)). This might be due to it's easy and fast setup and usage \autocite{sudarsana2019use}. Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).

@@

At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)). These peaks aligns with the near-total closures of schools due to covid 19 in Figure \ref{fig:covidImpactWorld}. The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects. Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}). The inability to reopen school fully due to the increasing Covid-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.

@@

The attention in the education sector is low compared to all category. However a significant peak can be observed during the first academic break after declaring Covid 19 as a pandemic. Similar peak can be observed in 2021 . But that significant compared to 2020 peak.

@@

In this study we investigated whether the Google trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions. This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during covid 19 pandemic. A few previous published studies \autocite{vaughan2014web,kansal2021google} have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.

>>>>

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
            pdfkeywords={Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

@@

\begin{document}
\maketitle
\begin{abstract}
Distance Education has a long history. However COVID-19 has made a new era of distance education. With the increasing demand, various different distance learning solutions have been introduced for different education purposes. In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions. Both visual and analytical approaches were used to analyse global level web search querie during COVID-19 pandemic. The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions. The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes. R code and data to reproduce the results of this work are available in the online supplementary materials.
\end{abstract}
\begin{keywords}
Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries
\end{keywords}

@@

COVID-19 is still one of the most important priorities of governments and media of many countries all around the world. Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}. As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}. In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.

@@

Many sectors have been affected ever since the outbreak of COVID-19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 \autocite{unesco2020covid}. With the start of the pandemic a near total closure of schools were observed all around the world. Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).

@@

Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}. Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}. Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.

The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as teachers, students and parents. Teachers were not fully aware of their obligations and how to connect with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID-19 pandemic. Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}. This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home. Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures. Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown. This situation left internet as the only medium to support all these educational processes.

According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught. In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure. There are many different distance learning solutions. Yet very little is known about the uptake of different solutions or about their effectiveness. Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and need of urgent responses.

Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting suitable distance education solutions. According to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as it can indicate the prominence of the product in the market, its usability and its impact. Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search. Therefore, the internet represents a great opportunity to learn about the popularity and public attention towards a product or service. It is also a way to narrow down the search space and thereby identify most suitable product and services that meets customer needs.

Google Trends is an open source web analytics tool that allows users to get access to Internet search data that has the ability to provide deep insights into population behavior \autocite{nuti2014use}. As there is still a struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Trends search quiries can be used as a proxy of the popularity and the public interest in different distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID-19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms during COVID-19 pandemic? (3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?

We primarily analyzed quantitative digital footprint data on the internet from December 2019 to August 2021. The resulted Google Trend footprint provides a fast first step guide to identify the most popular distance education tools available for different education purposes. It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching. Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.

Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal. Having the most appropriate tool is imperative to successful completion of a task at hand. The findings of this study provides an essential initial guidance to select the most appropriate tool for different distance education needs. Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.

@@

To identify the impact of COVID-19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO \autocite{unesco2020covid}. Data related to COVID-19 cases was retrieved from the R software \autocite{rsoftware} package \texttt{coronavirus} \autocite{coronavirusr}. The package pulls data from the Johns
Hopkins University Center for Systems Science and Engineering (JHU CCSE) data repository \autocite{dong2020interactive}.

In this work we also used Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions. We used weekly Google Analytics search index series of various distance learning solutions during COVID-19 pandemic. We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of COVID-19 pandemic. Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time. We also studied the cross correlation between weekly web searches and the weekly cases of COVID-19.

@@

Google Trends allows to search for up to five queries each time \autocite{vaughan2014web}. We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support. The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.

@@

Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the COVID-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education. We also evaluated the relationship between weekly web searches and the number of weekly global cases of COVID-19. Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of COVID-19 related search volume data and distance education related search volume data.

@@

Search volume index graphs for COVID-19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic. Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic. According to \textcite{moore2011learning}, ``distance education'' is the most renowned descriptor used when referencing distance learning. However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``never version'' and ``improved version'' of distance learning. The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout COVID-19 pandemic. Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period. COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases. A similar decline could be detected in distance education related terms. further, a second was observed only in education related terms after the WHO announcement of the COVID-19 pandemic. This could be due to the near total closures of schools happened with the declaration of COVID-19 as a pandemic.

@@

The pairwise comparison performed using Dynamic Time Warping algorithm. Under this approach the two time series are stretched or compressed locally in order to make one resemble the other as much as possible. The distance between the two series is then computed by summing the distances of individual aligned elements \autocite{giorgino2009computing}. Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from COVID-19 on distance education. Both pairwise comparison show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)). The normalized distance between the two series in the two comparisons are 0.1326197 and 0.1828299, respectively. This also provides further confirmation for the similarity of the google trend search profiles.

The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).

The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)). In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide. By the time it was know as ``corona'', it was not a global concern and was mostly limited to China. During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic all around the world. This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis.

Later with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020. With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}). With this sudden unexpected movement, online learning also became an equally important global concern, similar to COVID-19 pandemic. This similar behaviour in public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line from approximately 120 to 150 on the Query Index axis. This also confirms the impact of COVID-19 on distance education and online learning.

Then, a cross correlation analysis was performed to estimate the lag between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}). A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''. The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms. The significant cross correlation values at lag one or two also confirms the sudden impact of COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)). However, no significant synchronization was observed between the COVID-19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).

@@

\caption{Cross-correlation analysis between distance learning related online search terms and COVID-19 relate online search terms during COVID-19 pandemic}\label{fig:ccfAnalysis}
\end{figure}

@@

Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during COVID-19 pandemic (Figure \ref{fig:plot1} (a)). This might be due to it's easy and fast setup and usage \autocite{sudarsana2019use}. Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).

@@

At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)). These peaks aligns with the near-total closures of schools due to COVID-19 in Figure \ref{fig:covidImpactWorld}. The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects. Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}). The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.

@@

The attention in the education sector is low compared to all category. However a significant peak can be observed during the first academic break after declaring COVID-19 as a pandemic. Similar peak can be observed in 2021 . But that significant compared to 2020 peak.

@@

In this study we investigated whether the Google trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions. This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic. A few previous published studies \autocite{vaughan2014web,kansal2021google} have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
